\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-04-30 project-writings/paper.tex
Make a note about removing nobes
spellcheck

This approach does not work for us for two reasons: first, IonMonkey
does not use Cytron et al.'s algorithm for computing SSA form; it
instead takes advantage of the structured nature of Javascript code to
go to SSA form without computing dominance frontiers \footnote{More
  precisely, it takes advantage in an ad-hoc and terrifying way of
  poorly documented structured invariants and ``source notes'' in the
  bytecode that the SpiderMonkey frontend emits.}.  Furthermore, we
decided that instead of putting the program into XSA form at the
beginning and leaving it there through the compilation phase, it would
be a better idea to introduce $\beta$ nodes solely for performing
range analysis and remove them afterwards. This is for two main
reasons: first, as a question of engineering effort, we did not want
to teach every pass of the compiler how to deal with $\beta$ nodes,
and the IonMonkey team does not want us to change the IL in a pervasive
way simply for range analysis; second, $\beta$ nodes are copies that
are not allowed to be propagated away, which may inhibit optimization
passes (like global value numbering).

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.} .  Pseudocode for the algorithm is shown
in Figure \ref{fig:add_betas}. The algorithm traverses the control
flow group, looking for blocks that have been entered by a jump
conditioned on a relational operation. When we find a variable being
compared to a constant, we compute a range and use it to create a new
$\beta$ node. We want to propagate this range information to all uses
of the variable where the branch is guarenteed to have been taken;
this is accomplished by replacing all uses of original instruction
that are dominated by the new $\beta$ instruction with the $\beta$. To
do this, we take advantage of IonMonkey's use chains to efficiently
iterate over all uses of an instruction.

In order to do this, we needed to be able to efficiently determine
whether one block is dominated by another. The best way to do that
that existed in IonMonkey was to walk up the dominator tree looking
for the block being tested against. We felt that it would be too
inefficient to need to do this for every use of every variable tested
with a conditional. To do this, we annotated each basic block with its
index in a pre-order traversal of the dominator tree. Since IonMonkey
basic blocks already tracked how many other blocks they dominated, we
can now easily check whether a block dominates another by checking
whether the potential descendent's index falls within the range of the
ancestor's dominated blocks.

@@

>>>>

This approach does not work for us for two reasons: first, IonMonkey
does not use Cytron et al.'s algorithm for computing SSA form; it
instead takes advantage of the structured nature of Javascript code to
go to SSA form without computing dominance frontiers \footnote{More
  precisely, it takes advantage in an ad-hoc and terrifying way of
  poorly documented structured invariants and ``source notes'' in the
  bytecode that the SpiderMonkey front-end emits.}.  Furthermore, we
decided that instead of putting the program into XSA form at the
beginning and leaving it there through the compilation phase, it would
be a better idea to introduce $\beta$ nodes solely for performing
range analysis and remove them afterwards. This is for two main
reasons: first, as a question of engineering effort, we did not want
to teach every pass of the compiler how to deal with $\beta$ nodes,
and the IonMonkey team does not want us to change the IL in a pervasive
way simply for range analysis; second, $\beta$ nodes are copies that
are not allowed to be propagated away, which may inhibit optimization
passes (like global value numbering).

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.} .  Pseudocode for the algorithm is shown
in Figure \ref{fig:add_betas}. The algorithm traverses the control
flow group, looking for blocks that have been entered by a jump
conditioned on a relational operation. When we find a variable being
compared to a constant, we compute a range and use it to create a new
$\beta$ node. We want to propagate this range information to all uses
of the variable where the branch is guaranteed to have been taken;
this is accomplished by replacing all uses of original instruction
that are dominated by the new $\beta$ instruction with the $\beta$. To
do this, we take advantage of IonMonkey's use chains to efficiently
iterate over all uses of an instruction.

In order to do this, we needed to be able to efficiently determine
whether one block is dominated by another. The best way to do that
that existed in IonMonkey was to walk up the dominator tree looking
for the block being tested against. We felt that it would be too
inefficient to need to do this for every use of every variable tested
with a conditional. To do this, we annotated each basic block with its
index in a pre-order traversal of the dominator tree. Since IonMonkey
basic blocks already tracked how many other blocks they dominated, we
can now easily check whether a block dominates another by checking
whether the potential descendant's index falls within the range of the
ancestor's dominated blocks.

@@

The method for removing beta nodes is trivial. Traverse the program
removing beta nodes and replacing all uses of them with their
arguments. Since beta nodes only are placed at the beginning of basic
blocks, we only need to iterate through the $\beta$ node prefix of a
block.
